Learning to Group and Order AMR Semantics for General Purpose Discourse Planning

### 2015-12-10 paper/paper.tex
quickfix for figure labelling issue
revisions through discrete optimization

From a high level, our system is a pipeline, which takes an AMR representation
of the information set that is to be communicated as input, and outputs a
sequence of AMR representations, which are the AMRs representing the sentences
containing all of the information in the information set.

@@

For each subproblem, a learner was trained that could score the ``goodness" of
a given partitioning or ordering, depending on the problem. To find the best
possible partitioning or ordering, we use a greedy search optimization
procedure, which is described more fully in ``Discrete Optimization".

@@

Subgraph selection solves the problem of splitting an AMR graph into a set of
subgraphs which are optimal by some criterion, such that the union of the
subgraphs is the original graph. Said more plainly, the problem is to split the paragraph-level semantic graph 
into a set sentence-level semantic graphs.

There are $2^{|E|}$ subgraphs for a graph with $|E|$ edges. If we consider every set of $N$ subgraphs, we would have to consider $N^{2^{|E|}}$ sets of subgraphs! Obviously this is infeasible. Fortunately, we can exploit the natural language characteristics of a semantic graph to dramatically reduce the number of subgraph sets we have to consider.

First, we define a {\em fact} in our semantic graph as the subgraph formed by performing a graph-traversal on a root of the graph (that is, a node with no parents). An example is shown in figure \ref{fig:amr_example_3_merged_s1_highlighted}. By design, the roots of an AMR graph will be verb instances. Therefore, we can intuitively think of the rooted graph traversal as a verb + the required graph pieces to fulfill the verb's arguments. A fact, then, is the the smallest piece of a graph that could be used to fully specify the semantics of a sentence during generation. Let $F$ refer to the set of facts in our graph.

@@

\subsubsection{Training}
Our problem then is to find the best partition from the $B_{|F|}$ candidates. To do this, we trained a scoring function using a logistic regression model with class-weighted instances. We assumed that the partitioning given by the sentences in a paragraph was the correct partitioning, and gave it a label of $1$. This means that we had only positive examples in the training set, so we added negative examples by sampling $k$ other partitions from the set of partitions, giving each a label of $0$.

@@

We can test each component of the pipeline in isolation before combining them, to see how well each is performing. To test the Subgraph Selection scorer, I started with the 214 partitions in our test dataset (each with label 1), and sampled a random partition (with label 0) for a class-balanced test dataset with 428 instances. I generated features for each instance and ran the test set through the classifier, which later serves as a scoring function. The evaluation metrics are reported in table \ref{tab:subgraph_selection_eval}. The classifier performs very well in isolation; this is expected as it is just differentiating partitions between ``actual subgraph partition" and ``random subgraph partition''.

In addition to evaluating the test set, we can inspect the feature weights in Table \ref{tab:subgraph_selection_features} to get an idea of what the scoring function is doing. Looking at the first two features, we see that very small and very large paragraphs are penalized (since the ``number of sentences'' feature is positive but the square is negative). After that, we see the two most important features are the mean and standard deviation of the pairwise subgraph similarity; both features have large negative values. The large negative value on the mean pairwise subgraph similarity implies that the classifier prefers paragraphs for which similar facts are grouped into the same sentence, rather than spread throughout the paragraph. The negative value on the standard deviation implies that the classifier prefers a uniform rate of new information introduced in each sentence; there should be the same amount of information overlap between sentences.

@@

Once we have split a merged graph into a partition of $p$ sentences, there are
$s!$ ways to order the sentences. Ideally, there exists some ordering of the
sentences that communicates the ideas of the merged graph in the best possible
way. Subgraph ordering is the problem of finding this optimal ordering.

@@

Once again, we assumed that the ordering of sentences in the dataset was the
optimal ordering for that paragraph, so we needed to create non-optimal
examples to train on. This was done by randomly reordering the sentences for
$k=20$ reorderings for each paragraph.

We trained a Ridge regression model on the dataset. For the labels of the
training examples, we used the unnormalized Kendall's tau coefficient to
determine the distance between a given ordering and the optimal.

As noted in Related Work, our ordering model does not require the number of
sentences to be fixed to a particular value. If it were to impose such a
requirement, then the subgraph selection, which is further upstream in our
pipeline, may not be able to find its optimal partitioning simply because it
was a partitioning into a different number of sentences. As a result, our model
could not rely on any features that deal with particular sentences (e.g. the
similarity between the second and fifth sentences). So, we tried to use
features that would generalize to paragraphs of any length, but would still
vary with the ordering of those sentences within the paragraph, and all
features had to be summarized (mean, standard deviation, min, max) over the
paragraph. Features included the Jaccard similarity between adjancent
sentences, between sentences that are two sentences apart, and the Jaccard
similarity between the union of adjacent sentences and the overall paragraph.
The following table summarizes the features used for subgraph ordering:

\begin{table}
\centering
\begin{tabular}{|r|c|c|}
\hline
Feature & \\ \hline\hline
Pairwise Jaccard & Jaccard similarity of the AMR graphs for sentences $s_i,s_{i+1}$ \\ \hline
Two-apart Jaccard & Jaccard similarity of the AMR graphs for sentences $s_i,s_{i+2}$ \\ \hline
Pairwise node Jaccard & Jaccard similarity of just the vertex sets for sentences $s_i,s_{i+1}$ \\ \hline
Combined Jaccard & Jaccard similarity between the paragraph AMR and $s_i \cup
s_{i+1}$ \\ \hline
\end{tabular}
\caption{Subgraph Ordering Features}
\end{table}

@@

The mean, standard deviation, min and max for both the baseline and greedy
Kendall tau  are reported in the table below.

\begin{table}
\centering
\begin{tabular}{|r|c|c|c|c|c|}
\hline
Approach & Mean  & Std.  & Min & Max \\ \hline\hline
Baseline & 5.009 & 2.091 & 0   & 10  \\ \hline
Greedy   & 1.360 & 0.910 & 0   & 5   \\ \hline
\end{tabular}
\caption{Evaluation results for Subgraph Ordering}
\end{table}

@@

In both subproblems, we trained classifiers which assign a score to how ``good"
an instance is with regard to that problem. So, to find the best instance, we
could just enumerate all possible permutations or combinations of the input,
and select the instance which maximizes (or, in the case of subgraph ordering,
minimizes) the classifier's score. However, in the case of subgraph selection,
for a paragraph of $n$ facts, there are $O(2^n)$ possible partitions. In the
case of subgraph ordering, for a paragraph of $n$ sentences, there are $O(n!)$
possible orderings. Neither of these are very computationally efficient, so we
perform discrete optimization to approach an optimal solution, rather than
enumerating all possible inputs. It is worth noting that in practice, this is
more of a problem for the subgraph selection module, since the partitions
returned by the subgraph selector are generally not too large, so $n$ for the
subgraph orderer is not too bad.

For discrete optimization, we implemented a greedy search procedure as follows:
\begin{algorithm}
\caption{Greedy search procedure}
\label{greedy_search_alg}
\begin{algorithmic}
\STATE Input: $s$ // start state \\
\STATE Input: $classifier$ \\
\STATE $Q \gets [s]$ // queue of states to visit \\
\STATE $opt, optv, oldoptv \gets null, 0, 1$
\WHILE{$Q\neq\emptyset$ and $optv \neq oldoptv$}
    \STATE $S \gets \{s: s \in Q \}$
    \STATE $Q \gets \emptyset$
    \STATE $oldoptv \gets optv$
    \FORALL{$s \in S$}
        \STATE $N \gets neighbors(s)$\\
        \STATE Add all elements of $N$ to $Q$\\
        \IF{$classifier.score(s) > optv$}
            \STATE $opt, optv \gets s, classifier.score(s)$\\
        \ENDIF
    \ENDFOR
\ENDWHILE
\end{algorithmic}
\end{algorithm}

Additionally, we used a simulated annealing approach, but only for the
subgraph ordering problem.

>>>>

From a high level, our system is a pipeline, which takes as input an AMR representation
of all information that is to be communicated, and outputs a
sequence of AMR representations, which are the AMRs representing a sequence of sentence semantics that could be used to communicate the 
information set.

@@

For each subproblem, a scoring function was trained to evaluate the ``goodness" of
a given partitioning or ordering, depending on the problem. To find the best
possible partitioning or ordering, we use a greedy search optimization
procedure, which is described more fully in ``Discrete Optimization".

@@

Subgraph selection solves the problem of splitting an AMR graph into a set of
subgraphs which are optimal by some criterion, such that the union of the
subgraphs is the original graph. Said more plainly, the problem is to split the paragraph-level semantic graph into groups of information
which are sentence-level semantic graphs.

There are $2^{|E|}$ subgraphs for a paragraph graph with $|E|$ edges. If we consider every set of $N$ subgraphs, we would have to consider $N^{2^{|E|}}$ sets of subgraphs! Obviously this is infeasible. Fortunately, we can exploit the natural language characteristics of a semantic graph to dramatically reduce the number of subgraph sets we have to consider.

First, we define a {\em fact} in our semantic graph as the subgraph formed by performing a graph-traversal from a root of the graph (that is, a node with no parents). An example is shown in Figure \ref{fig:amr_example_3_merged_s1_highlighted}. By design, the roots of an AMR graph will be verb instances; therefore, we can intuitively think of the rooted graph traversal as a verb + the required graph pieces to fulfill the verb's arguments. Because a syntactically correct sentence requires a verb and its arguments, a {\em fact} is the the smallest piece of a graph that could be used to fully specify the semantics of a sentence. Let $F$ refer to the set of all facts in our graph.

@@

\subsubsection{Training}
Our problem then is to find the best partition from the $B_{|F|}$ candidates. To do this, we trained a scoring function using a logistic regression model with class-weighted instances. We assumed that the fact-partition given by the sentences of a paragraph was correct, and gave it a label of $1$. This means that we had only positive examples in the training set, so we added negative examples by sampling $k$ other valid partitions, giving each a label of $0$.

@@

To test the Subgraph Selection scorer, we started with the 214 partitions in our test dataset (each with label 1), and for each one sampled a random partition (with label 0), resulting in a class-balanced test dataset with 428 instances. Features were generated for each instance and the test set was passed through the classifier. The evaluation metrics are reported in table \ref{tab:subgraph_selection_eval}. The classifier performs very well in isolation; this is expected as it is differentiating partitions between ``actual subgraph partition" and ``random subgraph partition''.

In addition to evaluating the test set, we can inspect the feature weights in Table \ref{tab:subgraph_selection_features} to get an idea of what the scoring function is doing. Looking at the first two features, we see that very small and very large paragraphs are penalized (since the ``number of sentences'' feature is positive but the square is negative). After that, we see the two most important features are the mean and standard deviation of the pairwise subgraph similarity. Both features have large negative values. The large negative value on the mean pairwise subgraph similarity implies that the classifier prefers paragraphs for which similar facts are grouped into the same sentence, rather than spread throughout the paragraph. The negative value on the standard deviation implies that the classifier prefers a uniform rate of new information introduced in each sentence.

@@

Once we have split our semantic graph into a partition of $p$ sentence graphs, there are
$s!$ ways to order the sentences. Ideally, there exists some ordering of the
sentences that communicates the ideas of the merged graph in the best possible
way. Subgraph ordering is the problem of finding this optimal ordering.

@@

Once again, we assumed that the ordering of sentences in the dataset was the
optimal ordering for that paragraph, so we needed to create non-optimal
examples to train on as well. This was done by randomly generating
$k=20$ sentence reorderings for each paragraph.

We trained a Ridge regression model on the dataset to predict the Kendall Tau Distance of the ordering
(the distance between a given ordering and the optimal). This is in accordance with prior work, suggesting
the use of this metric \cite{lapata2006automatic}.

As noted in Related Work, our ordering model does not require the number of
sentences to be fixed to a particular value. If it were to impose such a
requirement, then the subgraph selection, which is further upstream in our
pipeline, may not be able to find its optimal partitioning simply because it
was a partitioning into a different number of sentences. As a result, our model
could not rely on any features that deal with particular sentences (e.g. the
similarity between the second and fifth sentences). So, we tried to use
features that would generalize to paragraphs of any length, but would still
vary with the ordering of those sentences within the paragraph. To ensure we 
had the same number of features for each instance, all features had to be summarized (mean, standard deviation, min, max) over the
paragraph. Features included the Jaccard similarity between adjancent
sentences, between sentences that are two sentences apart, and the Jaccard
similarity between the union of adjacent sentences and the overall paragraph.
Table \ref{tab:order_features} summarizes the features used for subgraph ordering.

\begin{table}
\centering
\label{tab:order_features}
\begin{tabular}{|r|c|c|}
\hline
Feature & \\ \hline\hline
Pairwise Jaccard & Jaccard similarity of the AMR graphs for sentences $s_i,s_{i+1}$ \\ \hline
Two-apart Jaccard & Jaccard similarity of the AMR graphs for sentences $s_i,s_{i+2}$ \\ \hline
Pairwise node Jaccard & Jaccard similarity of just the vertex sets for sentences $s_i,s_{i+1}$ \\ \hline
Combined Jaccard & Jaccard similarity between the paragraph AMR and $s_i \cup
s_{i+1}$ \\ \hline
\end{tabular}
\caption{Subgraph Ordering Features}
\end{table}

@@

The mean, standard deviation, min, and max for both the baseline and greedy
Kendall tau are reported in Table \ref{tab:order_eval}.

\begin{table}
\centering
\begin{tabular}{|r|c|c|c|c|c|}
\hline
Approach & Mean  & Std.  & Min & Max \\ \hline\hline
Baseline & 5.009 & 2.091 & 0   & 10  \\ \hline
Greedy   & 1.360 & 0.910 & 0   & 5   \\ \hline
\end{tabular}
\label{tab:order_eval}
\caption{Evaluation results for Subgraph Ordering}
\end{table}

@@

In both subproblems, we trained classifiers which assign a score to how ``good"
an instance is with regard to that problem. So, to find the best instance, we
could just enumerate all possible permutations or combinations of the input,
and select the instance which maximizes (or, in the case of subgraph ordering,
minimizes) the classifier's score. However, in both cases, there is an exponential number of possible orderings.
It is not computationally feasible to evaluate each, so we
perform discrete optimization to approach an optimal solution, rather than
enumerating all possible inputs. It is worth noting that in practice, this is
more of a problem for the subgraph selection module, since the number of partitions
returned by the subgraph selector is generally not too large.

For discrete optimization, we implemented a greedy search outlined in Algorithm \ref{greedy_search_alg}.

\begin{algorithm}
\caption{Greedy search procedure}
\label{greedy_search_alg}
\begin{algorithmic}
\STATE Input: $s$ // start state \\
\STATE Input: $classifier$ \\
\STATE $Q \gets [s]$ // queue of states to visit \\
\STATE $opt, optv, oldoptv \gets null, 0, 1$
\WHILE{$Q\neq\emptyset$ and $optv \neq oldoptv$}
    \STATE $S \gets \{s: s \in Q \}$
    \STATE $Q \gets \emptyset$
    \STATE $oldoptv \gets optv$
    \FORALL{$s \in S$}
        \STATE $N \gets neighbors(s)$\\
        \STATE Add all elements of $N$ to $Q$\\
        \IF{$classifier.score(s) > optv$}
            \STATE $opt, optv \gets s, classifier.score(s)$\\
        \ENDIF
    \ENDFOR
\ENDWHILE
\end{algorithmic}
\end{algorithm}

In addition to the greedy search, we experimented with simulated annealing for the
subgraph ordering problem.
